Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-16 ICML-submission/notions-of-linear-separability.tex
Fix section 3

\begin{definition}[Weak linear separability]
\label{definition:weak-linear-separability}
Let $(V,\ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, and $\gamma$ be a positive real number. Let $L = ((x_1, y_1), (x_2,
y_2), \dots, (x_T, y_T))$ be a set of labeled examples in $V \times \{1,2,\dots,K\}$.
\chicheng{This $L$ notion may be a bit confusing - feel free to edit}

@@

\begin{definition}[Strong linear separability]
\label{definition:strong-linear-separability}
Let $(V,\ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, and $\gamma$ be a positive real number. Let $L = ((x_1, y_1), (x_2,
y_2), \dots, (x_T, y_T))$ be labeled examples in $V \times \{1,2,\dots,K\}$.

@@

The notion of weak linear separability with a margin is standard. It is used in
the full-information setting to upper bound the number of mistakes of the
\textsc{Multiclass Perceptron} algorithm. It is a folklore result that if a set
of labeled examples is separable with a margin $\gamma$ and the norm of the
examples is bounded by $R$, then \textsc{Multiclass Perceptron} algorithm makes
at most $\left\lfloor 2(R/\gamma)^2 \right \rfloor$ mistakes. Another folklore
result is that \textsc{Multiclass Perceptron} is essentially optimal in the
sense that any deterministic algorithm must make $\frac 1 2 \left\lfloor (R/\gamma)^2
\right \rfloor$ mistakes in the worst case.
For completeness, we state the
\textsc{Multiclass Perceptron} algorithm and give proofs of both of these
results in Appendix~\ref{section:multiclass-perceptron-proofs}.

@@

The notion of strong linear separability has appeared in the literature; see
e.g.~\citet{Chen-Chen-Zhang-Chen-Zhang-2009}.
Intuitively, strong linear
separability means that, for each class $i$, the set of examples belonging to
class $i$ and the set of examples belonging to the remaining $K-1$ classes are
separated by a linear classifier $w_i$ by a margin $\frac \gamma 2$.

It is easy to see that if a set of labeled examples are strongly linearly
separable with margin $\gamma$, then it is also weakly linearly separable with
the same margin. Indeed, if $w_1, w_2, \dots, w_K \in V$
satisfy \eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2},
\eqref{equation:strong-linear-separability-3} then they satisfy
\eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2}.

In the special case of $K=2$, if a set of labeled examples are weakly linearly
separable with a margin $\gamma$, then it is also strongly linearly separable with
the same margin. Indeed, if $w_1, w_2$ satisfy
\eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2} then $w_1' = \frac{w_1 - w_2}{2}$,
$w_2' = \frac{w_2 - w_1}{2}$ satisfy
\eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2},
\eqref{equation:strong-linear-separability-3}.
Equation~\eqref{equation:strong-linear-separability-1} follows from
$\norm{w_i'}^2 \le (\frac{1}{2} \norm{w_1} + \frac{1}{2} \norm{w_2})^2 \le
\frac{1}{2}\norm{w_1}^2 + \frac{1}{2}\norm{w_2}^2 \le \frac{1}{2}$ for $i=1,2$.
Equations~\eqref{equation:strong-linear-separability-2} and~\eqref{equation:strong-linear-separability-3} follow from the fact that $w_1' - w_2' = w_1 - w_2$.

@@

While the examples are weakly linearly separable with a positive margin
$\gamma$, they are \emph{not} strongly linearly separable with any positive
margin $\gamma$. For instance, there do not exist linear separators that
separates the examples belonging to the gray class from the examples belonging
to the remaining two classes.
}
\label{figure:weakly-linearly-separable-examples-with-margin}
\end{figure}

>>>>

\begin{definition}[Weak linear separability]
\label{definition:weak-linear-separability}
Let $(V,\ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, and $\gamma$ be a positive real number. Let $L = ((x_1, y_1), (x_2,
y_2), \dots, (x_T, y_T))$ be a sequence of labeled examples in $V \times
\{1,2,\dots,K\}$.

@@

\begin{definition}[Strong linear separability]
\label{definition:strong-linear-separability}
Let $(V,\ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, and $\gamma$ be a positive real number. Let $L = ((x_1, y_1), (x_2,
y_2), \dots, (x_T, y_T))$ be a sequence of labeled examples in $V \times
\{1,2,\dots,K\}$.

@@

The notion of weak linear separability with a margin is standard. It is used in
the full-information setting to upper bound the number of mistakes of the
\textsc{Multiclass Perceptron} algorithm. It is a folklore result that if a set
of labeled examples is separable with a margin $\gamma$ and the norm of the
examples is bounded by $R$, then \textsc{Multiclass Perceptron} algorithm makes
at most $\left\lfloor 2(R/\gamma)^2 \right \rfloor$ mistakes. Another folklore
result is that \textsc{Multiclass Perceptron} is essentially optimal in the
sense that any (possibly randomized) algorithm must make $\frac{1}{2}
\left\lfloor (R/\gamma)^2 \right \rfloor$ mistakes in the worst case. For
completeness, we state the \textsc{Multiclass Perceptron} algorithm and give
proofs of both of these results in
Appendix~\ref{section:multiclass-perceptron-proofs} in the supplementary
material.

@@

The notion of strong linear separability has appeared in the literature; see
e.g.~\citet{Chen-Chen-Zhang-Chen-Zhang-2009}. Intuitively, strong linear
separability means that, for each class $i$, the set of examples belonging to
class $i$ and the set of examples belonging to the remaining $K-1$ classes are
separated by a linear classifier $w_i$ with margin $\frac{\gamma}{2}$.

It is easy to see that if a sequence of labeled examples is strongly linearly
separable with margin $\gamma$, then it is also weakly linearly separable with
the same margin. Indeed, if $w_1, w_2, \dots, w_K \in V$ satisfy
\eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2},
\eqref{equation:strong-linear-separability-3} then they satisfy
\eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2}.

In the special case of $K=2$, if a sequence of labeled examples is weakly
linearly separable with a margin $\gamma$, then it is also strongly linearly
separable with the same margin. Indeed, if $w_1, w_2$ satisfy
\eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2} then $w_1' = \frac{w_1 - w_2}{2}$,
$w_2' = \frac{w_2 - w_1}{2}$ satisfy
\eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2},
\eqref{equation:strong-linear-separability-3}.
Equation~\eqref{equation:strong-linear-separability-1} follows from
$\norm{w_i'}^2 \le (\frac{1}{2} \norm{w_1} + \frac{1}{2} \norm{w_2})^2 \le
\frac{1}{2}\norm{w_1}^2 + \frac{1}{2}\norm{w_2}^2 \le \frac{1}{2}$ for $i=1,2$.
Equations~\eqref{equation:strong-linear-separability-2}
and~\eqref{equation:strong-linear-separability-3} follow from the fact that
$w_1' - w_2' = w_1 - w_2$.

@@

While the examples are weakly linearly separable with a positive margin
$\gamma$, they are \emph{not} strongly linearly separable with any positive
margin $\gamma$. For instance, there does \emph{not} exist a linear separator
that separates the examples belonging to the gray class from the examples
belonging to the remaining two classes.
}
\label{figure:weakly-linearly-separable-examples-with-margin}
\end{figure}
